\PaperTitle

### 2020-05-17 paper/filter_bubbles.tex
fix the figures in the paper, add appendix, and rewrite section 4

\documentclass[format=acmsmall, review=false]{acmart}
\usepackage{acm-ec-20}
\usepackage{booktabs} % For formal tables
\usepackage[ruled]{algorithm2e} % For algorithms

@@

% Anonymized submission.
\author{Submission 20}

@@

\xhdr{Model} We consider a model of user choice that has the following components.
\par
The first component of our model is that users sequentially consume items and face large choice sets. In our setting of interest users are long-lived, but they only consume a small fraction of this choice set over their lifetime.
\par
The second component of our model is that users have \textit{uncertainty} about their valuation of the different items. This is motivated both by the fact that many markets where recommender systems are deployed contain experience goods, whose true value can only be learned after consumption, and the fact that such uncertainty is why recommender systems exist in the first place. Thus, users face a sequential decision-making problem under uncertainty.
\par 
The third, and key, component of our model is that consumption of an item reveals information that changes user beliefs about the valuation of similar items. Unlike in standard sequential decision-making problems, in our model once an item is consumed all uncertainty about its valuation is resolved. This valuation provides information for consumers that enables them to update their beliefs about similar items. This exploits the fact that the valuations of similar items are correlated which assists users in navigating the vast product space. The idea that users make similarity-based assessments to guide their choice has grounding in theoretical work in decision theory \cite{gilboa1995case} and empirical evidence of how users navigate large choice sets \cite{schulz2019structured}.
\par 
The fourth component of our model is that recommendation provides users with information about the true valuations. We model the realized valuations as being a weighted sum of a common-value and an idiosyncratic component. This formulation gives a stylized notion of predictability of user preferences where the idiosyncratic component is inherently unpredictable given other individuals' preferences and the common-value component is what the recommender can learn from previous users' data. We suppose that the recommender knows the common-value component for each item and combines it with individuals' beliefs over the product space when designing personalized recommendation.

@@

\section{Our Model and Preliminaries}
\subsection{Preliminaries on Expected Utility Theory}
\noindent For every item $n$ in the product space $\mathcal J$, we assume that each user $i$ assigns a monetary equivalent $x_{i,n} \in \mathbb R$ to the experience of consuming it. Each user can value the same item differently. However, we assume that users have the same utility over money, given by a utility function $u: \mathbb R \to \mathbb R$, strictly increasing and continuous. So, ex-post, the value of item $n$ for user $i$ is given by $u(x_{i,n})$ However, before consuming the item, the user does not know exactly how she will value it. In particular, even users that will end up having the same ex-post valuation of item $n$ may differ in their ex-ante valuation because they hold different beliefs about it. We denote by $p_{i}$ the beliefs item $i$ has about how she will value each of the items in the product space. Note that this implies that consuming item $n$ is the same as taking a gamble. Each user evaluates the item according to the associated expected utility associated with the item, i.e. $U_i(n)=\mathbb E_{p_i}[u(x_n)]$. 
\par

Risk aversion captures the how different users react to the risk associated to a particular consumption opportunity. It is formalized as follows: a given gamble $x$ takes real values and follows distribution $p$. Then, for every gamble $x$, there is a certain amount of money that makes the user indifferent between taking the gamble or taking the sure amount of money. This sure amount of money is called the \textit{certainty equivalent} of gamble $x$ and is denoted as $\delta(x)$. A user $i$ is more risk-averse than another user $j$ when for every gamble $x$, user $i$ requires a lower value to give up the gamble when compared to user $j$, i.e. $\delta_i(x)$. Therefore, a more risk-averse user is more willing to avoid the risk of taking the gamble. We assume that the utility function takes a flexible functional form $u(x)=1-\exp(-\gamma x)$ for $\gamma\ne0$ and $u(x)=x$ for $\gamma=0$ -- known as constant absolute risk-aversion preferences (from hereon CARA). Higher $\gamma$ implies higher risk-aversion with $\gamma=0$ corresponding to the risk-neutral case and $\gamma>0$ to the risk-averse one. Our formulations here follow standard economic decision-theory (see \cite{mas1995microeconomic} for a textbook treatment of these topics).
\par

@@

We will consider three cases. The case of primary interest is \textit{recommendation} where the recommender observes values accrued and knows $V$ but does not know neither $V_i$.\footnote{We do not consider the acquisition of information for the recommender to know $V$ and suppose that she has sufficient data to learn $V$ with arbitrary precision at $t = 0$.} However, the recommender does know the users' beliefs $\bar V_i$. Thus, at any given period, the recommender provides a personalized recommendation that combines the knowledge of the common value component $V$ with the user beliefs $\bar V_i$. Knowing the user's beliefs about her valuation of each item become crucial in this case: just providing the user with information about $V$ may change the user's original ranking, but, without considering the user's beliefs, she will not necessarily follow the recommendation.\footnote{The notion of recommendation that we consider is idealized where the recommendation does the Bayesian updating for users. As users face a very difficult and costly Bayesian updating instance, a useful recommendation system provides recommendations taking into account this aspect and combines the history of past consumption and value realizations with the user's beliefs. 
An alternative that is equivalent would be the recommender simply providing $V$ without knowing $V_i$. In this case, the recommendation will be that the user $i$ chooses $r_{i,t} \in \arg \max_{n\in \mathcal J\backslash C^T_i} \mathbb E[u(x_n) \mid C_i^T]$
but we assume the recommender provides full information about $V$. For instance, the recommender could display the whole distribution of utilities reported by other users or even its average, which is a good proxy for the common value component. It is possible for the recommender to only recommend the best item given such information, that is, the item with highest common value component. However, in that case, recommending only a single item generates costly Bayesian updating to the user and provides little guidance as to what she should indeed pick as the common value might be of little importance when compared to the idiosyncratic component. Therefore, we assume that the recommender reports the whole $V$, which results in higher expected welfare in choices and is not costly to implement.
The results would be equivalent, but this would shift the burden of Bayesian updating to users.}
\par

@@

We simulate over risk-aversion parameters $\gamma \in \{ 0, 0.3, 0.6, 1, 5 \}$, \ standard-deviation $\sigma \in \{ 0.25, 0.5, 1.0, 2.0, 4.0 \}$, correlation parameters $\rho\in \{ 0, 0.1, 0.3, 0.5, 0.7, 0.9 \} $ and degree of idiosyncrasy of individual valuations $\beta \in \{ 0, 0.4, 0.8, 1, 2, 5\}$. The range of considered parameter values cover the relevant portions of the parameter space in order to provide full insight into the behavior of the model. 
\par
When we consider results varying a single parameter, we group the results over the other parameters and provide reports varying only the parameter of interest. We report results for a relatively small consumption history $T=20$ with a product space size $N=200$. Unless otherwise reported, our results are robust to increasing $N$.
\par

\begin{figure}
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.9\linewidth]{figures/rho_pos_consumption_dist_N_200T_20_overall.pdf}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.9\linewidth]{figures/rho_zero_consumption_dist_N_200T_20_overall.pdf}
  \label{fig:sfig2}
\end{subfigure}
\caption{Consumption path with (Left) and without (Right) correlation between values}
\label{fig:correlation_consumption_path}
\end{figure}

@@

In sum, filter bubble effects can only arise when there is an inherent correlation between the realized utilities of the items in the product space. However, this is often the case in most environments where recommender systems are deployed. When there is a correlation between the realized utilities then filter bubbles can naturally arise due to the nature of how individuals acquire additional information about the remaining items. We have shown that, unless users are provided with additional information to guide their consumption choices, then these information spillovers and user risk-aversion can lead users into filter bubbles. When users consume high valued items, they exploit the underlying correlation across different products' values, stronger for similar products, which leads them to increasingly consume products in narrower and narrower portions of the product space. Risk aversion may lead users into performing local consumption even when they have a low valuation of nearby items just because they know what to expect from the item. Recommendation leads to these effects being mitigated by providing users with additional information on products outside the already explored portions of the product space. Further, if all uncertainty was resolved as in the oracle regime, then such behavior is not present.

@@

\begin{finding}\label{finding_welfare_gap}
The impact of recommendation on consumer welfare is as follows:
\begin{enumerate}
\item Under oracle recommendation, welfare is constant as $\rho$ increases.
\item Under no recommendation, welfare is increasing as $\rho$ increases.
\item Recommendation introduces welfare gains relative to no recommendation, but these gains are decreasing as $\rho$ increases. 
\end{enumerate}
\end{finding}
\par 
Finding~\ref{finding_welfare_gap} states our main findings of the impact of recommendation on ex-post welfare, which are drawn from Figure~\ref{fig:diversity_welfare_correlation}. The most interesting observation is that the value of recommendation decreases as $\rho$ decreases. The intuition for this is clear as recommendation provides users with information that allows them to better guide their decisions and increase welfare. However, as $\rho$ increases users get increasingly more information simply from consuming items since the realized utility is now more informative about the utility of nearby items. Thus, since consumption decisions themselves reveal information, the information that recommendation reveals provides less value.
\par 
Figure~\ref{fig:diversity_welfare_common_value} shows how diversity and welfare vary across recommendation regimes as the the strength of the common-value component, $\beta$, varies. As expected, when $\beta = 0$, these values are the same. However, as $\beta$ grows the idiosyncratic component of user valuations becomes increasingly less important and the recommendation regime begins to converge to the oracle regime. Indeed, we see that both product diversity and welfare increase under recommendation, whereas under no recommendation we see that there is a decrease in diversity with little changes to consumer welfare. This is intuitive as $\beta$ controls the informativeness of recommendation about user preferences.
\par 
One striking observation is that the decrease in diversity does not appear to be associated with a decline in welfare. Indeed, it appears that the opposite is the case - that low diversity is associated with higher welfare and vice versa. Thus, we next explore the relationship between welfare and diversity.
\begin{figure}
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/rho_diversity_N_200_T_20.pdf}
\end{subfigure}
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/rho_welfare_N_200_T_20.pdf}
\end{subfigure}
\caption{User Welfare and Diversity as the strength of correlation, $\rho$, varies}\label{fig:diversity_welfare_correlation}
\end{figure}

@@

Finding~\ref{finding_diversity_welfare_corr} summarizes our findings on the relationship between diversity and welfare. Figure \ref{fig:diversity_welfare_ra} shows how diversity and welfare correlate for the no recommendation case as we vary the degree of risk aversion. When there is no risk-aversion then there is a negative correlation between welfare and diversity. This is since, with no risk-aversion, in a given period a user will select the good that she currently believes has the highest expected value. High product diversity in this case can arise from a user who consumes an item she disliked and updates her beliefs about nearby items negatively. As a result, in the following period she will pick an item far away in the product space from the item that was previously consumed. If instead the user valued highly the item that she had consumed, then she is more likely to pick a nearby item. The information spillovers therefore lead to high product diversity being negatively correlated with welfare.
\par
This only happens since $\gamma = 0$ leads to users only caring about the expected value of the product. However, as we saw in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, increasing $\gamma$ can lead to lower diversity and increasingly local consumption due to the fact that the degree of uncertainty now impacts users' choices. This weakens the negative relationship between diversity and welfare since both negative and positive experiences with a good reduce uncertainty about surrounding items. This leads to the inverted-U shape found in Figure \ref{fig:diversity_welfare_ra} when $\gamma$ is relatively large (e.g $\gamma = 5$) though diversity and welfare are virtually uncorrelated in the data. In the recommendation and oracle regimes, under risk neutrality ($\gamma=0$), welfare and diversity and uncorrelated, while under risk aversion ($\gamma=5$), it is possible to observe in Figure \ref{fig:diversity_welfare_ra_partial} an actual positive relation between diversity and welfare as recommendations are able to reduce uncertainty and facilitate exploration of the product space.
\begin{figure}[t]
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_rn.pdf}
\end{subfigure}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_ra.pdf}
\end{subfigure}\\
\caption{Diversity vs. Welfare, \textbf{No-Recommendation}, $\gamma = 0$ (Left), Diversity vs. Welfare, $\gamma = 5$ (Right)}\label{fig:diversity_welfare_ra}
\end{figure}

\begin{figure}[t]
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_rn_partial.pdf}
\end{subfigure}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.05\linewidth]{figures/diversity_welfare_ra_partial.pdf}
\end{subfigure}\\
\caption{Diversity vs. Welfare, \textbf{Recommendation}, $\gamma = 0$ (Left), Diversity vs. Welfare, $\gamma = 5$ (Right)}\label{fig:diversity_welfare_ra_partial}
\end{figure}

@@

\section{Recommender System Evaluation}
In this section we discuss how the insights from our model of user decision-making can be used to inform the evaluation and design of recommender systems.
\par

The classic approach to recommender system evaluation is to predict user ratings for items and compare how accurate this prediction is to recorded ratings data, either explicitly given by users or inferred from behavioral data. Further, given this predicted rating, the recommender system should recommend to users the items with the highest predicted rating \cite{adomavicius2005toward}.
\par

There has been a recent movement away from such evaluation measures due to the observation that accurate recommendations are not necessarily useful recommendations \cite{mcnee2006being}. Our model illustrates why this could be the case. Consider the domain of movie recommendation and suppose a user has just watched the movie \textit{John Wick} and rated it highly. A recommender system attempting to predict user ratings may then predict that this user is very likely to enjoy the sequel, \textit{John Wick: Chapter Two}, as well. However, the user herself may also have made this inference since the two movies are very similar to each other and, after enjoying \textit{John Wick}, the user will update her beliefs about \textit{John Wick: Chapter Two}. Thus, if the recommender system recommends this movie to the user then this recommendation is not useful since the recommendation gives the user little information that they did not already know. The key insight is that the recommendation is not useful since \textit{it ignores the inference the user themselves made and their updated beliefs}. The user may watch \textit{John Wick: Chapter Two}, then, even without recommendation and the value of the recommendation was small.
\par

As a result, several alternative recommendation evaluation metrics, such as serendipity \cite{kotkov2016survey}, coverage \cite{ge2010beyond}, novelty \cite{vargas2011rank}, and many others have been proposed in order to develop recommender systems that produce recommendations that are more useful for users. We use our model to motivate an alternative recommendation evaluation metric that provides an alternative perspective on the serendipity metric proposed in the literature. Serendipitous recommendations are said to ``have the quality of being both unexpected and useful" \cite{maksai2015predicting} and we use this as a guide towards our definition.

Our approach to serendipity comes from the intuition that a recommendation is only useful insofar as it leads a user to take a better action than she would without recommendation. This approach stems from the idea that the value of recommendation is the marginal utility gain that a user receives from recommendation.\footnote{This notion is similar to the ``value of information" notion utilized in the literature in economics on the pricing of information goods \cite{bergemann2018design}.} This intuitively should come from providing information about products that, without additional information, users assign a low likelihood to being the best products currently available for them to consume and so would not consume them without these being recommended.
\par

This differs from other approaches in the literature since it depends on the beliefs of the users and what item the user would consume without recommendation. For instance, \cite{vargas2011rank, kaminskas2014measuring} propose unexpectedness metrics that look at the dissimilarity of the recommended items to what the recommender already knows the user likes either via a content-based approach or a collaborative-based approach. This depends only on the resulting item-set that the users get recommended and not necessarily user beliefs or the change in user action. \cite{adamopoulos2014unexpectedness} use an expected utility approach but use this to measure unexpectedness in terms of deviations from the ``expected consideration set" of a user and do not explicitly incorporate user beliefs.
\par

Our model allows us to give precise definitions for what it means for a recommendation to be \textit{unexpected} and \textit{useful}. Unexpected recommendations are given by recommended items that, given their beliefs, users have low expected utility from consuming, irregardless of what the true realized value is. Recall from Findings \ref{finding_local_consumption} and \ref{finding_diversity_welfare_corr} that, particularly if users are risk-averse, they will consume content in increasingly narrow portions of the product space and this may be detrimental to their realized welfare. As a result, identifying the portions of the product space where users may have significant uncertainty and utilizing recommendation to reduce this uncertainty can be welfare-enhancing for users. This is especially the case if this reduction of uncertainty leads users to explore under-explored portions of the product space where true realized utility is high, but expected utility, given beliefs, is low before recommendation.
\par

Useful recommendations are those that give users information that would change the item they consume relative to no recommendation and lead to an increase in expected utility given the updated beliefs when comparing the planned choice before and after the recommendation. In the previous example, recommending \textit{John Wick: Chapter Two} is not useful precisely because the expected utility before and after recommendation is the same since it does not change behavior. The most useful recommendations are those that (1) are informative given user beliefs and (2) induce users to consume unexpected goods.
\par

>>>>

\documentclass[format=acmsmall, review=false]{acmart}
\settopmatter{printacmref=false, printccs=false}
\usepackage{booktabs} % For formal tables
\usepackage[ruled]{algorithm2e} % For algorithms

@@

% Anonymized submission.

@@

\xhdr{Our Model} We consider a model of user choice that has the following components.
\par
The first component of our model is that users sequentially consume items and face large choice sets. In our setting of interest users are long-lived, but they only consume a small fraction of this choice set over their lifetime.
\par
The second component of our model is that users have \textit{uncertainty} about their valuation of the different items. This is motivated both by the fact that many markets where recommender systems are deployed contain experience goods, whose true value can only be learned after consumption, and the fact that such uncertainty is why recommender systems exist in the first place. Thus, users face a sequential decision-making problem under uncertainty.
\par 
The third, and key, component of our model is that consumption of an item reveals information that changes user beliefs about the valuation of similar items. Unlike in standard sequential decision-making problems, in our model once an item is consumed all uncertainty about its valuation is resolved. This valuation provides information for consumers that enables them to update their beliefs about similar items. This exploits the fact that the valuations of similar items are correlated which assists users in navigating the vast product space. The idea that users make similarity-based assessments to guide their choice has grounding in theoretical work in decision theory \cite{gilboa1995case} and empirical evidence of how users navigate large choice sets \cite{schulz2019structured}.
\par 
The fourth component of our model is that recommendation provides users with information about the true valuations. We model the realized valuations as being a weighted sum of a common-value and an idiosyncratic component. This formulation gives a stylized notion of predictability of user preferences where the idiosyncratic component is inherently unpredictable given other individuals' preferences and the common-value component is what the recommender can learn from previous users' data. We suppose that the recommender knows the common-value component for each item and combines it with individuals' beliefs over the product space when designing personalized recommendation.

@@

\section{Our Model and Preliminaries}
\subsection{Preliminaries on Expected Utility Theory}
\noindent For every item $n$ in the product space $\mathcal J$, we assume that each user $i$ assigns a monetary equivalent $x_{i,n} \in \mathbb R$ to the experience of consuming it. Each user can value the same item differently. However, we assume that users have the same utility over money, given by a utility function $u: \mathbb R \to \mathbb R$, strictly increasing and continuous. So, ex-post, the value of item $n$ for user $i$ is given by $u(x_{i,n})$ However, before consuming the item, the user does not know exactly how she will value it. In particular, even users that will end up having the same ex-post valuation of item $n$ may differ in their ex-ante valuation because they hold different beliefs about it. We denote by $p_{i}$ the beliefs user $i$ has about how she will value each of the items in the product space. Note that this implies that consuming item $n$ is the same as taking a gamble. Each user evaluates the item according to the associated expected utility associated with the item, i.e. $U_i(n)=\mathbb E_{p_i}[u(x_n)]$. 
\par

Risk aversion captures how different users react to the risk associated to a particular consumption opportunity. It is formalized as follows: a given gamble $x$ takes real values and follows distribution $p$. Then, for every gamble $x$, there is a certain amount of money that makes the user indifferent between taking the gamble or taking the sure amount of money. This sure amount of money is called the \textit{certainty equivalent} of gamble $x$ and is denoted as $\delta(x)$. A user $i$ is more risk-averse than another user $j$ when for every gamble $x$, user $i$ requires a lower value to give up the gamble when compared to user $j$, i.e. $\delta_i(x)$. Therefore, a more risk-averse user is more willing to avoid the risk of taking the gamble. We assume that the utility function takes a flexible functional form $u(x)=1-\exp(-\gamma x)$ for $\gamma\ne0$ and $u(x)=x$ for $\gamma=0$ -- known as constant absolute risk-aversion preferences (from hereon CARA). Higher $\gamma$ implies higher risk-aversion with $\gamma=0$ corresponding to the risk-neutral case and $\gamma>0$ to the risk-averse one. Our formulations here follow standard economic decision-theory (see \cite{mas1995microeconomic} for a textbook treatment of these topics).
\par

@@

We will consider three cases. The case of primary interest is \textit{recommendation} where the recommender observes values accrued and knows $V$ but does not know $V_i$.\footnote{We do not consider the acquisition of information for the recommender to know $V$ and suppose that she has sufficient data to learn $V$ with arbitrary precision at $t = 0$.} However, the recommender does know the users' beliefs $\bar V_i$. Thus, at any given period, the recommender provides a personalized recommendation that combines the knowledge of the common value component $V$ with the user beliefs $\bar V_i$. Knowing the user's beliefs about her valuation of each item become crucial in this case: just providing the user with information about $V$ may change the user's original ranking, but, without considering the user's beliefs, she will not necessarily follow the recommendation.\footnote{The notion of recommendation that we consider is idealized where the recommendation does the Bayesian updating for users. As users face a very difficult and costly Bayesian updating instance, a useful recommendation system provides recommendations taking into account this aspect and combines the history of past consumption and value realizations with the user's beliefs. 
An alternative that is equivalent would be the recommender simply providing $V$ without knowing $V_i$. In this case, the recommendation will be that the user $i$ chooses $r_{i,t} \in \arg \max_{n\in \mathcal J\backslash C^T_i} \mathbb E[u(x_n) \mid C_i^T]$
but we assume the recommender provides full information about $V$. For instance, the recommender could display the whole distribution of utilities reported by other users or even its average, which is a good proxy for the common value component. It is possible for the recommender to only recommend the best item given such information, that is, the item with highest common value component. However, in that case, recommending only a single item generates costly Bayesian updating to the user and provides little guidance as to what she should indeed pick as the common value might be of little importance when compared to the idiosyncratic component. Therefore, we assume that the recommender reports the whole $V$, which results in higher expected welfare in choices and is not costly to implement.
The results would be equivalent, but this would shift the burden of Bayesian updating to users.}
\par

@@

We simulate over risk-aversion parameters $\gamma \in \{ 0, 0.3, 0.6, 1, 5 \}$, \ standard-deviation $\sigma \in \{ 0.25, 0.5, 1.0, 2.0, 4.0 \}$, correlation parameters $\rho\in \{ 0, 0.1, 0.3, 0.5, 0.7, 0.9 \} $ and degree of idiosyncrasy of individual valuations $\beta \in \{ 0, 0.4, 0.8, 1, 2, 5\}$. The range of considered parameter values cover the relevant portions of the parameter space in order to provide full insight into the behavior of the model. 
\par
When we consider results varying a single parameter, we group the results over the other parameters and provide reports varying only the parameter of interest. We report results for a relatively small consumption history $T=20$ with a product space size $N=200$.
\par

\begin{figure}
\caption{Local Consumption and Correlation, $N = 200$}
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.9\linewidth]{figures/rho_pos_consumption_dist_N_200T_20_overall.pdf}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.9\linewidth]{figures/rho_zero_consumption_dist_N_200T_20_overall.pdf}
  \label{fig:sfig2}
\end{subfigure}
\caption*{\scriptsize Notes: The figure shows the consecutive consumption path difference between the no recommendation, recommendation, and oracle regime. The figure on the left displays the average consecutive consumption distance aggregating over simulations where $\rho = 0$ and the figure on the right displays the average consecutive consumption distance aggregating over simulations where $\rho > 0$}
\label{fig:correlation_consumption_path}
\end{figure}

@@

In sum, filter bubble effects only arise when there is an inherent correlation between the realized utilities of the items in the product space. When there is a correlation between the realized utilities then filter bubbles can naturally arise due to the nature of how individuals acquire additional information about the remaining items. We have shown that, unless users are provided with additional information to guide their consumption choices, then these information spillovers and user risk-aversion can lead users into filter bubbles. When users consume high valued items, they exploit the underlying correlation across different products' values, stronger for similar products, which leads them to increasingly consume products in narrower and narrower portions of the product space. Risk aversion may lead users into performing local consumption even when they have a low valuation of nearby items just because they know what to expect from the item. Recommendation leads to these effects being mitigated by providing users with additional information on products outside the already explored portions of the product space. Further, if all uncertainty was resolved as in the oracle regime, then such behavior is not present.

@@

\begin{finding}\label{finding_welfare_gap}
The impact of recommendation on consumer welfare is as follows:
\begin{enumerate}
\item Under oracle recommendation, welfare is constant as $\rho$ increases.
\item Under no recommendation, welfare is increasing as $\rho$ increases.
\item Recommendation introduces welfare gains relative to no recommendation, but these gains are decreasing as $\rho$ increases. 
\end{enumerate}
\end{finding}
\par 
Finding~\ref{finding_welfare_gap} states our main findings of the impact of recommendation on ex-post welfare, which are drawn from Figure~\ref{fig:diversity_welfare_correlation}. The most interesting observation is that the value of recommendation decreases as $\rho$ decreases. The intuition for this is clear as recommendation provides users with information that allows them to better guide their decisions and increase welfare. However, as $\rho$ increases users get increasingly more information simply from consuming items since the realized utility is now more informative about the utility of nearby items. Thus, since consumption decisions themselves reveal information, the information that recommendation reveals provides less value.
\par 
Figure~\ref{fig:diversity_welfare_common_value} shows how diversity and welfare vary across recommendation regimes as the the strength of the common-value component, $\beta$, varies. As expected, when $\beta = 0$, these values are the same. However, as $\beta$ grows the idiosyncratic component of user valuations becomes increasingly less important and the recommendation regime begins to converge to the oracle regime. Indeed, we see that both product diversity and welfare increase under recommendation, whereas under no recommendation we see that there is a decrease in diversity with little changes to consumer welfare. This is intuitive as $\beta$ controls the informativeness of recommendation about user preferences.
\par 
One striking observation is that the decrease in diversity does not appear to be associated with a decline in welfare. Indeed, it appears that the opposite is the case - that low diversity is associated with higher welfare and vice versa. Thus, we next explore the relationship between welfare and diversity.
\begin{figure}
\caption{Relationship between User Welfare and Diversity, varying $\rho$}\label{fig:diversity_welfare_correlation}
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/rho_diversity_N_200_T_20.pdf}
\end{subfigure}
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/rho_welfare_N_200_T_20.pdf}
\end{subfigure}
\caption*{\scriptsize Notes: The figure on the left displays the relationship between $rho$, the correlation between valuations of items, and overall consumed product diversity. The figure on the right displays the relationship between $\rho$ and overall welfare.}
\end{figure}

@@

Finding~\ref{finding_diversity_welfare_corr} summarizes our findings on the relationship between diversity and welfare. Figure \ref{fig:diversity_welfare_ra} shows how diversity and welfare correlate for the no recommendation case as we vary the degree of risk aversion. When there is no risk-aversion then there is a negative correlation between welfare and diversity. This is since, with no risk-aversion, in a given period a user will select the good that she currently believes has the highest expected value. High product diversity in this case can arise from a user who consumes an item she disliked and updates her beliefs about nearby items negatively. As a result, in the following period she will pick an item far away in the product space from the item that was previously consumed. If instead the user valued highly the item that she had consumed, then she is more likely to pick a nearby item. The information spillovers therefore lead to high product diversity being negatively correlated with welfare.
\par
This only happens since $\gamma = 0$ leads to users only caring about the expected value of the product. However, as we saw in Findings \ref{finding_local_consumption} and \ref{finding_diversity}, increasing $\gamma$ can lead to lower diversity and increasingly local consumption due to the fact that the degree of uncertainty now impacts users' choices. This weakens the negative relationship between diversity and welfare since both negative and positive experiences with a good reduce uncertainty about surrounding items. This leads to the inverted-U shape found in Figure \ref{fig:diversity_welfare_ra} when $\gamma$ is relatively large (e.g $\gamma = 5$) though diversity and welfare are virtually uncorrelated in the data. In the recommendation and oracle regimes, under risk neutrality ($\gamma=0$), welfare and diversity and uncorrelated, while under risk aversion ($\gamma=5$), it is possible to observe in Figure \ref{fig:diversity_welfare_ra_partial} an actual positive relation between diversity and welfare as recommendations are able to reduce uncertainty and facilitate exploration of the product space.
\begin{figure}[t]
\caption{Diversity vs. Welfare, \textbf{No Recommendation}}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_rn.pdf}
\end{subfigure}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_ra.pdf}
\end{subfigure}\\
\caption*{\scriptsize Notes: The figure plots the relationship between diversity and welfare under the no recommendation regime. The figure on the left displays this relationship for when $\gamma = 0$ and the figure on the right for when $\gamma = 5$.}\label{fig:diversity_welfare_ra}
\end{figure}

\begin{figure}[t]
\caption{Diversity vs. Welfare, \textbf{Recommendation}}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_rn_partial.pdf}
\end{subfigure}
\begin{subfigure}{.45\textwidth}
\includegraphics[width=1.0\linewidth]{figures/diversity_welfare_ra_partial.pdf}
\end{subfigure}\\
\caption*{\scriptsize Notes: The figure plots the relationship between diversity and welfare under the recommendation regime. The figure on the left displays this relationship for when $\gamma = 0$ and the figure on the right for when $\gamma = 5$.}\label{fig:diversity_welfare_ra_partial}
\end{figure}

@@

\section{Recommender System Evaluation}
In this section we discuss how the insights from our model of user decision-making can inform the evaluation and design of recommender systems. The classic approach to recommender system evaluation is predicting user ratings for items and comparing how accurate this prediction is to recorded ratings data, either explicitly given by users or inferred from behavioral data. The recommender system should then recommend to users the items with the highest predicted rating \cite{adomavicius2005toward}.
There has been a recent movement away from such evaluation measures due to the observation that accurate recommendations are not necessarily useful recommendations \cite{mcnee2006being}. Our model illustrates a mechanism behind this observation. Consider the domain of movie recommendation and suppose a user has just watched the movie \textit{John Wick} and rated it highly. A recommender system attempting to predict user ratings may then predict that this user is very likely to enjoy the sequel, \textit{John Wick: Chapter Two}, as well. However, the user herself may also have made this inference since the two movies are very similar to each other. Thus, if the recommender system recommends this movie to the user then this recommendation is not useful since the recommendation gives the user little information that they did not already know. The key insight is that the recommendation is not useful since \textit{it ignores the inference the user themselves made and their updated beliefs}. The user may watch \textit{John Wick: Chapter Two}, then, even without recommendation, and the value of the recommendation was small.
This intuition implies that recommender systems should collect additional data beyond the type of data that is traditionally recorded. The first and most crucial type of data to collect is individual user \textit{beliefs} about items that they have not yet consumed. As illustrated by our model, these beliefs are what drive the future consumption decisions of users and understanding these beliefs is crucial for determining the value of recommending certain items.\footnote{Additionally, user beliefs contain information that may not have been observed by the recommender that only observes user choices on the platform. There are many sources of information that affect a user's beliefs and choices that are unobservable by recommender systems such as ratings, reviews, and friends. Other consumption choices, e.g. movies seen in the cinema, change beliefs about for instance how good a director is, which affect beliefs about other movies and guide choices on a streaming platform that is unable to observe these data. However, this information can be inferred by the platform by collecting both beliefs and consumption choices.} The second type of data that is critical for recommender system designers to collect is how user beliefs change over time and, in particular, not just how individuals value the item they just consumed, but also how it impacts their beliefs about the quality of similar items. The third type of data is the risk-aversion levels of users as our model illustrates that the risk preferences of users are important for understanding what information recommender systems can provide that materially leads users to alter their consumption patterns.
A natural follow-up question is how this additional data should be utilized in the design of good recommendations. Our model posits that recommendation provides value to users by providing them with information about the true valuation of an item if they were to consume it. Thus, the prediction problem for the recommender becomes predicting what item would the user choose with no recommendation and, correspondingly, what would be the most useful information to provide to the user that would lead her to consume a \textit{better} item than she would without recommendation. This links back to the intuition our model provided for the \textit{John Wick} example whereby collecting user beliefs and measuring how the user updated beliefs about similar goods would lead the recommender to understand that the user would consume \textit{John Wick: Chapter Two}. Our approach would therefore imply that, with this as a starting point, the recommender's problem would be to predict what is the most useful information to give the user leading them to change the item that they eventually consume.
There have been a number of alternative recommendation evaluation metrics proposed in the literature with the aim of providing more useful recommendations than those provided by accuracy metrics, such as serendipity \cite{kotkov2016survey}, coverage \cite{ge2010beyond}, novelty \cite{vargas2011rank}, and many others. Our approach most closely follows the set of proposed serendipity measures which are surveyed in \cite{kotkov2016survey}. As discussed by \cite{maksai2015predicting},  serendipitous recommendations are said to ``have the quality of being both unexpected and useful" which is in line with the primary intuition behind our approach. The primary difference between our proposed approach and those existing in the literature is that ours crucially hinges on understanding user beliefs and the risk-preferences of users. For instance, \cite{vargas2011rank, kaminskas2014measuring} propose unexpectedness metrics that look at the dissimilarity of the proposed recommended items compared to what the recommender already knows the user likes - either via a content-based approach or a collaborative-based approach. This metric depends only on the resulting item-set that the recommender proposes and not necessarily on the user's beliefs or how such a recommendation will change the item that the user consumes. \cite{adamopoulos2015unexpectedness} propose an expected utility approach but use this to measure unexpectedness in terms of deviations from the ``expected consideration set" of a user and do not explicitly incorporate user beliefs.
Indeed, our approach allows us to give precise definitions for what it means for a recommendation to be \textit{unexpected} and \textit{useful} in the spirit of serendipitous recommendations. Our evaluation measure leads to useful recommendations since it leads users towards better items than they would consume without recommendation. It further results in 	``unexpected" recommendations since it explicitly incorporates user beliefs and thus allows the recommender to understand how ``unexpected" a recommendation would be from the perspective of a user. Thus, a good recommendation provides users with information that leads them to consume an item that they would not have expected to consume without the information from the recommender and that they enjoyed consuming more than the item they would have without recommendation.
% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}
